### 2012-12-17 article.tex

\documentclass[a4paper,12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{fullpage,url}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{listings}
\usepackage{amssymb,amsmath,amsthm}
\usepackage{color}

\title{How to make a delicious pizza}
\author{Yuto Takei \\ The University of Tokyo}

\begin{document}

\maketitle

\begin{abstract}
  This is the abstract.
\end{abstract}

\section{Introduction}

* Formal verification
--> Dependent typing of SHP
--> Horn clauses

* Some
Background for Horn clause solving
* Dependent type determination

Definition of Horn clause solving (Informal)

Advantages
* Solution space consideration

* Why do we limit to integer space?

\section{Algorithm Overview}

\subsection{Preliminaries}

$ <num> ::= ... | -2 | -1 | 0 | 1 | 2 | ... $ .
$ <var> ::= x | y | ... $
$ <term> ::= <num> \cdot <var> $
$ <expr> ::=  <term> + <term> + \cdots + <num> >= 0 $
$ <la> ::= <expr>
         | <la> \wedge <la> $
$ <pvar> ::= A | B | ... $
$ <pred> ::= <pvar> ( <var>, <var>, ... ) $
$ <hornTerm> ::= <pvar> | <pred> $
$ <horn> ::= <hornTerm> \wedge <hornTerm> \wedge \cdots \longrightarrow
         <hornTerm> $

* We consider inequalities.
** NOTE: Even we have =, <>, >, <=, we can normalize to >=
         over integer space

\subsection{Generating a tree}

By receiving Horn clauses as an input for the algorithm, we build a
implication tree.

Def <Implication Tree>:

Implication tree is such a tree that every node is labeled by logical
atoms. The label of every node is implied by the conjunction of its
children's label.

* Renaming?

\subsection{Application of Farkas' Lemma}

Correct statement of Farkas' Lemma
Relationship to Craig's Interpolation.

\subsection{Solution space representation}

\section{Extensions}

\subsection{Extension to DAG structure}

* Building graph

* Splitting a graph into tree

* Merging with the same-name predicate conjunctively

\subsection{Linear arithmetics with disjunction}

In the previous section, we assumed that the input inequalities are
conjunctively connected.

* Why do we need this extension? *

We extend our inequalities to include disjunctions in the linear
equations.

$ <la> ::= <expr>
         | <la> \wedge <la>
         | <la> \vee <la> $

* Predicates on the direct path from the leaf with disjunction to the
root are connected by disjunction
* Other predicates are connected by conjunction

* Why do we process LA-disjunction before DAG?
** DAG to LA-Disjunction are invalid???

\subsection{Merging of pred-pred}

* Simplification between clauses

\subsection{General Horn clauses with Loops}

No limitation Horn clause solving

* Merging pred-pred between looped same name

\section{Experiment}
\section{Future Work}
\section{Conclusion}

\bibliographystyle{plain}
\bibliography{./biblio}

\end{document}