### 2013-08-26 article/experiment.tex

\chapter{Experiment}
\label{chap:experiment}

We incorporated our algorithm to the software model checker for
higher-order programs (MoCHi) \cite{conf/pldi/KobayashiSU11}.  It
is specifically designed to verify the correctness of higher-orde
programs written in a small subset of ML.  It infers dependent types
for input higher-order programs \cite{conf/ppdp/UnnoK09}.

Our implementation make use of Z3 \cite{conf/tacas/MouraB08} as
an SMT solver.

For simplicity of the implementation, we assume that the linear
inequalities in the integer space.  Because of this, our algorithm
lost the completeness.  By using Motzkin's transpose theorem
\cite{journals/networks/Rajan90}, it is possible to handle linear
inequalities in both $\mathbf{ax} + b \leq 0$ and $\mathbf{ax} + b <
0$.

\section{Interpolating theorem prover}

Our interpolating theorem prover is used as a sub-procedure to solve
Horn clauses.

\begin{itemize}
\item CSIsat \cite{conf/cav/BeyerZM08}
\item FOCI \cite{website/foci}
\item CLP-Prover \cite{website/clp}
\end{itemize}

\section{Horn clause solver}