### 2013-08-26 article/future.tex

\chapter{Future Work}
\label{chap:future}

\paragraph{Non-linear extension}
\cite{conf/cav/DaiXZ13}