### 2013-08-26 article/intro.tex

\chapter{Introduction}

Automated theorem proving plays an important role in the modern
program verification.  Particularly in the context of verification
with model checking techniques, an automated theorem prover is used to
abstract an infinite set of program states into a finite one so that
the model checker can virtually explore all possible program execution
paths.

One of the most common methods of such abstraction is predicate
abstraction\cite{conf/cav/GrafS97} within CounterExample-Guided
Abstraction Refinement (CEGAR) framework
\cite{conf/cav/ClarkeGJLV00,conf/popl/BallR02,conf/popl/HenzingerJMS02}.
An automated theorem prover finds appropriate predicates for every
program location.  An execution states at those locations is then
expressed by a Boolean valuation of the given predicates.  If the
abstraction is too coarse and a model checker discovers an infeasible
error execution path in an abstract system, the prover computes an
additional predicate at those program locations to refine the
abstraction predicates.

In order to obtain an additional abstraction predicate at program
locations along the infeasible counterexample, one class of automated
theorem provers, called an interpolating theorem prover
\cite{journals/tcs/McMillan05,conf/vmcai/RybalchenkoS07}, computes a
separating predicate called an interpolant between two sets of
constraints; the one consists of the constraints from the program
entry point to the location, and the other consists of the constraints
from the location to the program failure.  Those constraints include
variable assignments and assertions to be satisfied originated from
conditional branches.

The computed predicate by an interpolating theorem prover is satisfied
at the location under the execution in the concrete model along the
discovered counterexample, and any execution paths which satisfies the
predicate does not lead to the failure state.  After the refinement,
the infeasible path is ruled out and no longer found by the model
checker.

However, existing interpolating theorem provers may return too complex
a solution which is heavily affected by the constraints of the
specific path.  It is not favorable in terms of the program
verification because a complex predicate may not be sufficient to rule
out similar spurious paths in programs with loops and recursion calls.
This causes the model checker to discover spurious paths infinitely
which pass the same program location by loops or recursion calls at
different numbers of iteration, and the verification process does not
to terminate.

To avoid this, it is desired to discover an invariant formula for the
predicate abstraction.  Based on the hypothesis that program's
invariants tend to be simple formulas, we propose a new interpolating
algorithm which tries to minimize the number of conjunctions and
disjunctions in a solution.  The predicate obtained by this algorithm
may become similar to the invariant, and the number of trials in model
checking may be reduced.

Additionally, we extend our interpolating algorithm to solve a
symmetric interpolation problem, in which the algorithm computes
predicates at multiple program locations along a path at once.  This
enables the abstraction predicates at multiple locations to be updated
at the same computation.

Finally, we extend our algorithm to solve Horn clause solving
problems.  It allows us to obtain predicates which are highly likely
to be program invariants, by computing the same abstraction predicates
inside loops and recursion calls on different iterations, and by
unifying predicates at the same location over multiple paths.  The
problems are generated by encoding constraints over execution paths
with unknown predicates which describe functions' constraints between
parameters and return values.

The rest of the paper is structured as follows.
Chapter~\ref{chap:interpolation} proposes a new interpolating
algorithm to obtain simple solutions, and extends it to solve
symmetric interpolation problems.  Chapter~\ref{chap:horn} describes a
new Horn clause solving algorithm by extending the previous
interpolating algorithm.  Chapter~\ref{chap:experiment} describes
experiments and illustrates the result to confirm the effect of our
proposed algorithm.  Chapter~\ref{chap:future} gives a discussion of
our improvements and proposes further improvements to make.
Chapter~\ref{chap:related} mentions related work done by others.
Finally, Chapter~\ref{chap:conclusion} reviews the impact of our
research and brings the conclusion.